---
title: "Data Exploration of Bacteria in Latrines in Vietnam and Tanzania"
author: "Group 11: Freja Dahl Hede, Maika Jensen, Malene Nørregaard Nielsen, <br>Sofie Betzer Rossen, Sofie Theisen Honoré"
date: May 12th 2021
output: ioslides_presentation
---

```{r setup, include=FALSE}
# Clear workspace ---------------------------------------------------------
rm(list = ls())
knitr::opts_chunk$set(echo = FALSE)

# Load libraries ----------------------------------------------------------
library(ggplot2)
library(knitr)
library(kableExtra)
library(tidyverse)
```

```{r, include = FALSE}
#### 1. Prepare Data for printing ####
my_data_raw_counts <- read_csv(
  file = "/cloud/project/data/_raw/All_Good_P2_C03-1.csv")
my_data_raw_taxonomy <- read_csv(
  file = "/cloud/project/data/_raw/All_Good_P2_C03_Taxonomy-1.csv")
my_data_raw_ENV <- read_csv(
  file = "/cloud/project/data/_raw/ENV_pitlatrine.csv")
my_data_clean_aug <- read_tsv(
  file = "/cloud/project/data/03_my_data_clean_aug.tsv")
```

<!-- Introduction -->
# Introduction 

## Introduction 
- Explorative data analysis of bacterial latrine data from Tanzania and Vietnam
- Torondel, B. et al., (2016), 
"Assessment of the influence of intrinsic environmental and geographical 
factors on the bacterial ecology of pit latrines",
Microbial Technology

<!-- Materials-->
# Materials & Methods

## Data 
3 sets of raw data

- Abundance data with count per OTU cluster per sample
- Taxonomy data per OTU cluster
- Environmental metadata per sample

## Raw Data Extraction
```{r}
# Random extract of data. Seed is set to 7 to ensure non-0 rows
set.seed(4)
my_data_raw_counts %>%
  select(1:10) %>%
  slice_sample(n = 10) %>%
  kable("html") %>%
  kable_styling(font_size = 12)
```

## Reference Data & Metadata
```{r, out.width='90%'}
set.seed(11)
# Extraxt of taxonomy data
my_data_raw_taxonomy %>%
  select(1:6) %>%
  slice_sample(n = 5) %>%
  kable("html") %>%
  kable_styling(font_size = 12)

# Extract of meta data
my_data_raw_ENV %>%
  slice_head(n = 5) %>%
  kable("html") %>%
  kable_styling(font_size = 12)
```


## Representation of Wrangling Process
```{r, out.width = '90%'}
knitr::include_graphics(
  path = '/cloud/project/doc/example_pivotlonger.png')
```

## Representation of Wrangling Process
```{r, out.width = '90%'}
knitr::include_graphics(
  path = '/cloud/project/doc/example_leftjoin.png')
```

## Clean and Augmented Data
```{r}
my_data_clean_aug %>%
  select(Samples, Class, site, everything()) %>%
  slice_head(n = 10) %>%
  select(1:10) %>%
  kable("html") %>%
  kable_styling(font_size = 9)
```

## Methods
- Hierarchical Clustering
- Pearson Correlations
- Principal Component Analysis
- KMeans Clustering
- Welch T-test

## Representation of Data Flow
```{r, out.width = '90%'}
knitr::include_graphics(
  path = '/cloud/project/doc/Flowchart_overview.png')
```

# Results
## Summary of Data
```{r, include = FALSE}
# Extract information from data
number_bacteria <- my_data_clean_aug %>%
  distinct(Class) %>%
  nrow()

number_countries <- my_data_clean_aug %>%
  distinct(site) %>%
  nrow()

number_samples <- my_data_clean_aug %>%
  distinct(Samples) %>%
  nrow()

number_samples_Tanzania <- my_data_clean_aug %>%
  filter(site == "Tanzania") %>%
  distinct(Samples) %>%
  nrow()

number_samples_Vietnam <- my_data_clean_aug %>%
  filter(site == "Vietnam") %>%
  distinct(Samples) %>%
  nrow()

max_OTU <- my_data_clean_aug %>%
  summarise(max(OTU_Count)) %>% 
  first()

min_OTU <- my_data_clean_aug %>%
  summarise(min(OTU_Count)) %>% 
  first()
```
- Number of bacteria classes: `r number_bacteria` (Including Unclassified)
- Number of sites: `r number_countries`
- Number of samples: `r number_samples`
    - Tanzania: `r number_samples_Tanzania`
    - Vietnam: `r number_samples_Vietnam`
- OTU Count:
    - Maximum in one sample: `r format(max_OTU, digits = 5)`
    - Minimum in one sample: `r format(min_OTU, digits = 5)`


## Initial Exploratory Analysis
```{r, out.width = '100%'}
knitr::include_graphics(
  path = '/cloud/project/results/TempCarboNH4pH_violinplot.png')
```

## Initial Exploratory Analysis
```{r, out.width = '100%'}
knitr::include_graphics(
  path = '/cloud/project/results/correlation_attributest.png')
```

## Initial Exploratory Analysis
```{r, out.width = '100%'}
knitr::include_graphics(
  path = '/cloud/project/results/heatmap_with_dendrogram.png')
```

## Initial Exploratory Analysis
```{r, out.width = '100%'}
knitr::include_graphics(
  path = '/cloud/project/results/heatmap_with_totalbar.png')
```

## Principal Components Analysis
```{r, out.width = '100%'}
knitr::include_graphics(
  path = '/cloud/project/results/pca_variance_plot.png')
```

## Principal Components Analysis
```{r, out.width = '100%'}
knitr::include_graphics(
  path = '/cloud/project/results/pca_plot.png')
```

## KMeans Analysis
```{r, out.width = '100%'}
knitr::include_graphics(
  path = '/cloud/project/results/elbowplot.png')
```

## KMeans Analysis
```{r, out.width = '100%'}
knitr::include_graphics(
  path = '/cloud/project/results/clust_4_pca.png')
```

## Difference of Bacteria Abundance between Countries
```{r, out.width = '90%'}
knitr::include_graphics(
  path = '/cloud/project/results/Manhattan_plot.png')
```

## Difference of Bacteria Abundance between Countries
```{r, out.width = '80%'}
knitr::include_graphics(
  path = '/cloud/project/results/class_boxplots.png')
```

## Difference of Bacteria Abundance between Countries
```{r, out.width = '100%'}
knitr::include_graphics(
  path = '/cloud/project/results/correlation_attributes_bacteria.png')
```

## Difference of Bacteria Abundance between Countries
```{r, out.width = '100%'}
knitr::include_graphics(
  path = '/cloud/project/results/gammaproteobacteria_pH_barplot.png')
```

## Difference of Bacteria Abundance between Countries
```{r, out.width = '100%'}
knitr::include_graphics(
  path = '/cloud/project/results/gammaproteobacteria_Temp_barplot.png')
```

## Difference of Bacteria Abundance between Countries
```{r, out.width = '100%'}
knitr::include_graphics(
  path = '/cloud/project/results/gammaproteobacteria_Carbo_barplot.png')
```

## Difference of Bacteria Abundance between Countries
```{r, out.width = '100%'}
knitr::include_graphics(
  path = '/cloud/project/results/gammaproteobacteria_NH4_barplot.png')
```

# Discussion

## Discussion
- Difference between OTU count between countries

- Influence from environmental factors

- Hypothesis based on the exploratory analysis
